---
title: "Lexical Richness as a Lens: Exploring the Influence of Social Class on Authors' Writings"
author: 
  - Qi Er (Emma) Teng
  - Wentao Sun
  - Yang Cheng 
  
thanks: "Code and data are available at: https://github.com/dwz92/Analyzing-the-impact-of-social-class-on-authors-writings."
date: today
date-format: long
abstract: "In this study, we explore the influence of social class on lexical richness in the works of authors  after the Great Depression, utilizing a comprehensive analysis of lexical data from literary works and social context of authors. Our findings reveal upper-class authors exhibiting a higher Corrected Type-Token Ratio which indicates greater lexical richness, and reflecting their social milieu. This study employs a thorough lexical analysis, aiming to elucidate the complex interplay between class background and language use in literature. This research contributes to our understanding of how social class background shapes language use in literature, offering insights into the broader cultural forces that drive linguistic creativity."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# install.packages("tidyverse")
# install.packages("dplyr")
# install.packages("here")
# install.packages("mapdata")
# install.packages("ggplot2")
# install.packages("haven")
# install.packages("kableExtra")

library(tidyverse)
library(dplyr)
library(here)
library(ggplot2)
library(modelsummary)
library(kableExtra)
```

# Introduction {#sec-intro}
The Great Depression was a major economic event of the 20th century that had a profound effect on the global economic situation and class divisions. The question of whether the writing styles of different authors are affected by their class of origin and the general economic climate deserves an in-depth discussion. The purpose of this paper is to gain a deeper understanding of how people lived and thought after the Great Depression by examining whether upper-class writers had a richer writing vocabulary to utilize in their essays than middle-class writers.


Utilizing a combination of quantitative data analysis and authorship studies, we investigated the vocabulary richness of texts used by upper-class and middle-class authors after the Great Depression. Our findings suggest that works produced by upper-class authors during the period 1930-1990 will have a richer vocabulary compared to middle-class writers. This phenomenon suggests that different classes of writers have different writing styles, and that economic downturns and international instability have had an impact on writers' writing styles. It also had a profound effect on the writers' personal lives, including education, life and thought. These influences, brought about by the broader context of the times, affect not only the choice of vocabulary used in the essay, but also the expression of the central idea of the whole essay, which in turn affects the minds of many civilian readers.


The paper is structured to facilitate a comprehensive understanding of the study and its implications. Following @sec-intro, @sec-data presents the data, detailing the data sources, analytical techniques, and the rationale behind the chosen methods. @sec-result discusses the results, elaborating on the observed trends and patterns in the authors writings. @sec-discussion provides an in-depth discussion of these findings, exploring potential factors influencing these trends, drawing connections to broader socio-economic issues, and providing suggestions for future research in this area.

# Estimand
In this study, our main object of estimation is the causal effect of different classes on the vocabulary use of writers' writing, i.e., changes that can be attributed to the writers' own class of origin, divorced from the economic factors of their times. This involves analyzing the richness of the writing vocabulary of authors from both the upper and middle classes, in order to understand how differences in education, life, and personal perceptions brought about by different class origins affect writers' ideas and writing styles.

# Data {#sec-data}

# TODO: explain

## Source and Methodology

# TODO: explain

## Variables

#TODO: explain raw text, sample raw text

```{r}
#| label: tbl-display-raw-data
#| tbl-cap: First Ten Row of Waste No tear by Hugh Garners
#| echo: false

waste_no_tears_t <- readLines(here::here("data/analysis_data/waste_no_tear_raw.txt"), encoding = "UTF-8")

waste_no_tears_t |>
  kable(
    col.names = c("Text"),
    booktabs = TRUE
  )

```
@tbl-display-raw-data, created with `kableExtra` [@citekableExtra], showcases the first ten rows of 
#TODO: Explain raw data




# TODO: explain For better visualization

```{r}
#| label: tbl-authors-cttr
#| tbl-cap: First Ten Rows of Authors' Corrected Type Token Ratio
#| echo: false

cleaned_data <- read.csv(here::here("data/analysis_data/cttr_all.csv"))

cleaned_data |>
  select(author, title, corrected_type_token_ratio) |>
  slice(1:10) |>
  kable(
    col.names = c("author", "title", "corrected type token ratio"),
    digits = 1,
    booktabs = TRUE
  )
```
@tbl-authors-cttr
# TODO: Explain 

## Measurements

#TODO: explainnnn

#Model
# TODO: Explain neg binom model

## Model set-up

# TODO: formula

## Model Justification

# TODO: expectation and why



# Results {#sec-result}

# Expalin
@sec-result 

## Hugh Garners CTTR and Length
```{r}
#| label: fig-trends-in-birth-rate
#| fig-cap: Trend in Birth Rates in US
#| echo: false


```


## David Herbert Lawrence CTTR and Length
```{r}
#| label: fig-trends-in-birth-rate-in-young-people
#| fig-cap: Trends in Birth Rates by Age Group
#| echo: false


```






## Nancy Mitford CTTR and Length
```{r}
#| label: fig-birth-rate-comparison-old-and-young
#| fig-cap: Birth Rate between Younger and Older Population
#| echo: false

```


## Powell Anthony CTTR and Length
```{r}
#| label: fig-change-in-birth-rates-by-state
#| fig-cap: Change in Birth Rates by State, 2004–2008 to 2015–2019
#| echo: false


```


\newpage

# Discussion {#sec-discussion}

The findings of this study offer a layered perspective on the decline in birth rates among young people in the United States, particularly in the context of the Great Recession. Our analysis provides an understanding of the multifaceted influences on demographic trends.

## Findings




## Economic Impact Insights



## Societal and Technological Influences



## Weaknesses and Future Research Directions
One limitation of our study is the potential influence of unobserved variables that may affect the lexical richness in authors' works, such as individual educational background or specific life experiences, which were not fully captured in our analysis. 
Moreover, while we used an improved TTR formula, its inherent sensitivity to text length variability and the possibility of other underlying linguistic factors were not entirely addressed. These could include stylistic choices or genre-specific language use that our study does not explore in depth.
Future research should aim to unravel these complex interactions further, perhaps by integrating more granular data on authors' personal histories or by employing a wider range of linguistic analysis tools. Additionally, exploring the application of advanced computational methods, such as natural language processing algorithms, could offer more detailed insights into how social class impacts lexical choice and style in literature. Understanding these nuanced relationships is vital for comprehensively interpreting the intersection of social factors and literary expression.




\newpage

\appendix

# Appendix {-}

# Data Manipulation and Cleaning {#sec-appendix}

# TODO: explainn





\newpage


# References
